Scottish Child Poverty Briefing Plots

### 2024-06-18 minos/outcomes/ChildPovReduction_Combined.Rmd
Added the income quintile plots based on the together summaries

---
title: "Scottish Child Poverty Briefing Plots"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: true
    df_print: paged
    code_folding: hide
  pdf_document:
    toc: true
---

@@

@@

  plot_data <- data %>%
    select(run_id, SF_12, hh_income, year, scenario) %>%
    pivot_wider(names_from = scenario,
                values_from = c(SF_12, hh_income),
                names_sep = '.') %>%
    mutate(across(starts_with("SF_12"), ~ . - SF_12.baseline),
           across(starts_with("hh_income"), ~ . - hh_income.baseline)) %>%
    rename_with(~ paste0("diff_", .), starts_with("SF_12")) %>%
    rename_with(~ paste0("diff_", .), starts_with("hh_income")) %>%
    pivot_longer(cols = starts_with("diff"),
                 names_to = c(".value", "scenario"),
                 names_sep = "\\.") %>%
    filter(scenario != 'baseline')

@@

We use an equation derived from table 4 (below) in [this paper by Lawrence and Fleishman](https://pubmed.ncbi.nlm.nih.gov/15090102/) to calculate QALYs from SF-12 MCS and PCS.

@@

@@

```{r}
SF12_change_income_quint(summary_group = 'whole_pop',
                                    scen.list = scen.list.rel)
```

@@

```{r}
SF12_change_income_quint(summary_group = 'families',
                                    scen.list = scen.list.rel)
```

@@

```{r}
SF12_change_simple(summary_group = 'treated_relative',
                   scen.list = scen.list.rel)

@@

```{r}
SF12_change_income_quint(summary_group = 'whole_pop',
                                    scen.list = scen.list.abs)
```

@@

```{r}
SF12_change_income_quint(summary_group = 'families',
                                    scen.list = scen.list.abs)
```

@@

```{r}
SF12_change_simple(summary_group = 'treated_absolute',
                   scen.list = scen.list.abs)

>>>>

---
title: "Scottish Child Poverty Briefing Plots"
output:
  pdf_document:
    toc: true
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: true
    df_print: paged
    code_folding: hide
---

@@

  print(p1)
  print(p2)
}

SF12_change_income_quint2 <- function(summary_group, scen) {

  summary_group <- paste0(summary_group, '_income_quint_together')
  scen <- paste0(scen, '_together')

  plot_data <- data %>%
    select(run_id, SF_12, income_quintile, year, scenario) %>%
    pivot_wider(names_from = scenario,
                values_from = SF_12,
                names_sep = '.',
                values_fn = list(SF_12 = function(x) x[1])) %>%
    mutate(diff_SF12 = intervention - baseline) %>%
    pivot_longer(cols = starts_with("diff"),
                 names_to = c(".value", "scenario"),
                 names_sep = "\\.") %>%
    select(run_id, income_quintile, year, diff_SF12)

  plot_data_confint <- plot_data %>%
      group_by(year, income_quintile) %>%
      summarise(n = n(),
                mean_SF12_change = mean(diff_SF12),
                margin_SF12 = qt(0.975, df = n - 1) * (sd(diff_SF12) / sqrt(n)),
                lower_SF12 = mean_SF12_change - margin_SF12,
                upper_SF12 = mean_SF12_change + margin_SF12)

  plot_data_confint$income_quintile <- factor(plot_data_confint$income_quintile)

  p2 <- ggplot(plot_data_confint, aes(x = year, y = mean_SF12_change, group = income_quintile, color = income_quintile, fill = income_quintile)) +
      geom_ribbon(aes(ymin = lower_SF12, ymax = upper_SF12), linetype = 'dotted', alpha = 0.3) +
      geom_line(linetype = 'dashed') +
      geom_point() +
      geom_hline(yintercept = 0, linetype = 'dashed') +
      geom_vline(xintercept = 2030, linetype = 'dotted', alpha = 0.5) +
      labs(title = 'Treatment on the Treated',
           subtitle = 'SF_12') +
      xlab('Year') +
      ylab('Change in mean SF-12')

  #print(p1)
  print(p2)
}
```

@@

  plot_data <- data %>%
    select(run_id, SF_12, hh_income, year, scenario) %>%
    pivot_wider(names_from = scenario,
                values_from = c(SF_12, hh_income),
                names_sep = '.') %>%
    unnest(cols = c(starts_with("SF_12"), starts_with("hh_income"))) %>%
    mutate(across(starts_with("SF_12"), ~ . - SF_12.baseline),
           across(starts_with("hh_income"), ~ . - hh_income.baseline)) %>%
    rename_with(~ paste0("diff_", .), starts_with("SF_12")) %>%
    rename_with(~ paste0("diff_", .), starts_with("hh_income")) %>%
    pivot_longer(cols = starts_with("diff"),
                 names_to = c(".value", "scenario"),
                 names_sep = "\\.") %>%
    filter(scenario != 'baseline')

@@

We use an equation derived from table 4 (below) in [this paper by Lawrence and Fleishman](https://pubmed.ncbi.nlm.nih.gov/15090102/) to calculate QALYs from SF-12 Mental Component Score (MCS) and Physical Component Score (PCS).

@@

This rough calculation shows that a 0.1 absolute change in mean SF_12 across the Scottish population would equal over 5000 additional QALYs in a single year. These improvements can also accumulate over time. If we were able to achieve a 0.1 improvement in mean MCS for Scotland in year 1, and maintain this 0.1 improvement for 10 years, the resulting improvement in QALYs would be 56,000+.

NOTE: This calculation assumes that PCS is unchanged when MCS can vary. This is not realistic, but the calculation is still useful to provide context for the results to follow.

@@

```{r}
# SF12_change_income_quint(summary_group = 'whole_pop',
#                                     scen.list = scen.list.rel)
SF12_change_income_quint2(summary_group = 'whole_pop',
                                    scen = 'ChildPovertyReductionRELATIVE_2')
```

@@

```{r}
# SF12_change_income_quint(summary_group = 'families',
#                                     scen.list = scen.list.rel)
SF12_change_income_quint2(summary_group = 'families',
                                    scen = 'ChildPovertyReductionRELATIVE_2')
```

@@

```{r}
# SF12_change_simple(summary_group = 'treated_relative',
#                    scen.list = scen.list.rel)

@@

```{r}
# SF12_change_income_quint(summary_group = 'whole_pop',
#                                     scen.list = scen.list.abs)
SF12_change_income_quint2(summary_group = 'whole_pop',
                                    scen = 'ChildPovertyReductionABSOLUTE')
```

@@

```{r}
# SF12_change_income_quint(summary_group = 'families',
#                                     scen.list = scen.list.abs)
SF12_change_income_quint2(summary_group = 'families',
                                    scen = 'ChildPovertyReductionABSOLUTE')
```

@@

```{r}
# SF12_change_simple(summary_group = 'treated_absolute',
#                    scen.list = scen.list.abs)
